\documentclass{article}
\usepackage[left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fancyhdr}
\usepackage{verbatim}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{pdfpages}
\usepackage{cancel}
\usepackage{physics}

\lstset{
    language=C++,
    basicstyle=\ttfamily\footnotesize,
    keywordstyle=\color{blue},
    commentstyle=\color{green},
    stringstyle=\color{red},
    numbers=left,
    numberstyle=\tiny\color{gray},
    frame=single,
    breaklines=true,
    captionpos=b,
}


\title{MTH 553: Lecture \# 14}
\author{Cliff Sun}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{one minute paper}[theorem]{One Minute Paper}

\pagestyle{fancy}
\lhead{\textbf{\thepage}\ \ \nouppercase{\rightmark}}
\chead{MTH 553: Lecture \# 14}
\rhead{Cliff Sun}

\begin{document}

\maketitle

\section*{Last time}

From last time, for the wave equation in three dimensions, we claim that the Kirchhoff formula solves 

\begin{equation}
    u(x,t) = \partial_t \left( t\int_{S^2} g(x + ct\xi)\frac{d\xi}{4\pi}\right) + t\int_{S^2}h(x + ct\xi)\frac{d\xi}{4\pi}
\end{equation}

Remarks:

\begin{enumerate}
    \item It involves the derivatives of $g$ as well as $g$. (unlike d'Alembert)
    \item Domain of dependence for $(x,t)$ is the sphere with center $x$ and radius $ct$. This means that as time progresses, more and more parts of $S^2$ influence $x$. This lives in the $x$ space. 
    \item Range of influence for $\tilde{x}$ is a surface of a cone $$|x - \tilde{x}| = ct$$, which is a union of spheres with radius $ct$. This lives in the x-t plane.
    \item Waves have finite propagation speed $c$
    \item Signals are sharp in 3-dim since the range of influence is a 2 dimension surface at each time.   
\end{enumerate}

\section*{HWE in 2 dimensions}

$u \in C^2(\mathbb{R}^2 \times [0,\infty))$ solves 

\begin{align*}
    u_{tt} &= c^2\nabla^2 u \\
    u(x,0) &= g(x_1,x_2) \\
    u_t(x,0) &= h(x_1,x_2)
\end{align*}

iff 

\begin{equation}
    u(x,t) = \frac{1}{4\pi}\frac{\partial}{\partial t}\left( 2t\int_{B^2}g(x + cty)\frac{dy}{\sqrt{1 - |y|^2}} \right) + \frac{1}{4\pi}2t\int_{B^2}h(x + cty)\frac{dy}{\sqrt{1 - |y|^2}}
\end{equation}

Where $B^2$ is the unit disk in $\mathbb{R}^2$. 

\begin{proof}
    Use Kirchhoff's formula. $u$ solves HWE in three dimensions, just not depending on $x_3$. Then take 
    \begin{equation}
        U(x_1,x_2,x_3,t) = u(x_1,x_2,t)
    \end{equation}
    \begin{equation}
        G(x_1,x_2,x_3) = g(x_1,x_2)
    \end{equation}
    \begin{equation}
        H(x_1,x_2,x_3) = h(x_1,x_2)
    \end{equation}
    Where $U$ satisfies the wave equation in three dimensions. Now, apply Kirchhoff. In the upper hemisphere of $S^2$, consider a point on the disk $B^2$. Now, projecting that onto 
    $S^2$ corresponds to a unit vector $\xi = (y_1,y_2, \sqrt{1 - |y|^2})$. Then $d\xi$ is the area of the graph of $z = \sqrt{1 - |y|^2}$. Then 
    \begin{equation}
        d\xi = \sqrt{1 + \left( \frac{\partial z}{\partial y_1} \right)^2 + \left( \frac{\partial z}{\partial y_2} \right)^2}
    \end{equation} 
    \begin{equation}
        = \frac{dy}{\sqrt{1 - |y|^2}}
    \end{equation}
    Here, this is the 3d projected area element of our 2d problem. Also, then 
    \begin{equation}
        x + ct \xi = (x_1 + ct y_1, x_2 + ct y_2, x_3 + ct\sqrt{1 - |y|^2})
    \end{equation}
    Finally, the sphere has two hemispheres so we collect a factor of $2$ (upper and lower are identical by symmetry). Therefore, we obtain equation (2).
\end{proof}

\subsection*{Remarks}

\begin{enumerate}
    \item This is called Hadamard's method of descent for solving 2-d WE using Kirchhoff's formula
    \item Domain of dependence for $(x,t)$ is a disk $B^2(x,ct)$ (disk centered at $x$ with radius $ct$). What points of $x$ influenced this point? 
    \item Range of influence, solid cone going positively. Namely
    \begin{equation}
        | x- \tilde{x}| \leq ct
    \end{equation}
    This is a union of disks with radius of $ct$. Note, this is because in the domain of dependence in 2-d is a circle, but in 3-d is the boundary of a ball (also called sphere)
    \item Propagation speed $c$
    \item No sharp signals (meaning signals live every in $B^2$), but they do decay in time. 
\end{enumerate}

\section*{HWE in $\mathbb{R}^n$, $n \geq 4$}

If $n$ is even, WE behaves like 3-d with sharp signals. If $n$ is odd, then WE behaves like 2-d with non-sharp signals. This is called Huygen's principle. 

\subsection*{NHWE in $\mathbb{R}^n$}

APply Duhamel to handle the forcing term. Use Kirchhoff, Hadamard as appropriate depending on dimensionality to compute $Z(x,t-s;s)$. 

\section*{Next time - Conservation of Energy}

\textbf{Example: n=1} $u_{tt} = c^2u_{xx}$. Assume $u(x,t)$ has compact support for each $t$. Then multiply the WE by $u_t$ and integrate with respect to $x$.

\begin{align*}
    0 &= \int_{-\infty}^{\infty}u_t(u_{tt} - c^2u_{xx})dx \\
    &= \int_{-\infty}^{\infty}\left( u_t u_{tt} + c^2 u_xu_{xt} \right) dx \text{ by parts} \\
    &= \frac{d}{dt}\frac{1}{2}\int_{-\infty}^{\infty}\left( u_t^2 + c^2 u_{x}^2 \right)dx
\end{align*}

This is the conservation of energy. 

\end{document}